---
title: "Pre-Class Work 02"
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE, cache = FALSE}
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
```

Part 1

1. Generate 200 random values from the standard exponential distribution and store them in a vector exp.draws.1. Find the mean and standard deviation of exp.draws.1.
```{r}
exp.draws.1 <- rexp(n=200) 
mean1 <- mean(exp.draws.1)
sd1 <- sd(exp.draws.1)
```
The mean of exp.draws.1 is `r mean1`.
The standard deviation of exp.draws.1 is `r sd1`.

2. Repeat, but change the rate to 0.2, 5, 7.3 and 10, storing the results in vectors called exp.draws.0.2, exp.draws.5, exp.draws.7.3 and exp.draws.10.
```{r}
exp.draws.0.2 <- rexp(n=200, rate=0.2) 
mean0.2 <- mean(exp.draws.0.2)
sd0.2 <- sd(exp.draws.0.2)
```
The mean of exp.draws.0.2 is `r mean0.2`.
The standard deviation of exp.draws.0.2 is `r sd0.2`.
```{r}
exp.draws.5 <- rexp(n=200, rate=5) 
mean5 <- mean(exp.draws.5)
sd5 <- sd(exp.draws.5)
```
The mean of exp.draws.5 is `r mean5`.
The standard deviation of exp.draws.5 is `r sd5`.
```{r}
exp.draws.7.3 <- rexp(n=200, rate=7.3) 
mean7.3 <- mean(exp.draws.7.3)
sd7.3 <- sd(exp.draws.7.3)
```
The mean of exp.draws.7.3 is `r mean7.3`.
The standard deviation of exp.draws.7.3 is `r sd7.3`.
```{r}
exp.draws.10 <- rexp(n=200, rate=10) 
mean10 <- mean(exp.draws.10)
sd10 <- sd(exp.draws.10)
```
The mean of exp.draws.10 is `r mean10`.
The standard deviation of exp.draws.10 is `r sd10`.

3. The function plot() is the generic function in R for the visual display of data. hist() is a function that takes in and bins data as a side effect. To use this function, we must first specify what we'd like to plot. a. Use the hist() function to produce a histogram of your standard exponential distribution. b. Use plot() with this vector to display the random values from your standard distribution in order. c. Now, use plot() with two arguments -- any two of your other stored random value vectors -- to create a scatterplot of the two vectors against each other.
```{r}
#a 
hist(exp.draws.1, main = "Histogram of exp.draws.1")

#b
sort(exp.draws.1, decreasing=FALSE) #sorting random values from standard distribution in increasing order 
plot(exp.draws.1, main = "Scatterplot of Standard Exponential Distribution in Order ")

#c
plot(exp.draws.5, exp.draws.10, main = "Scatterplot of Exponential Distributions with Rate 5 against 10")
```

4. We'd now like to compare the properties of each of our vectors. Begin by creating a vector of the means of each of our five distributions in the order we created them and saving this to a variable name of your choice. Using this and other similar vectors, create the following scatterplots and explain in words what is going on: a. The five means versus the five rates used to generate the distribution. b. The standard deviations versus the rates. c. The means versus the standard deviations.For each plot, explain in words what's going on.
```{r}
means <- c(mean1, mean0.2, mean5, mean7.3, mean10) #means of each of our five distributions in order we created them

#a
rates <- c(1, 0.2 ,5 ,7.3 ,10)
plot(means,rates, main = "Scatterplot of Means vs. Rates")
```

This is a scatterplot of the means of the five distributions on the x-axis versus the rates of the five distribution on the y-axis. This plot shows that as the means increase, the rates decrease. It is an exponentially distributed graph. 

```{r}
#b
sds <- c(sd1, sd0.2, sd5, sd7.3, sd10)
plot(sds, rates, main = "Scatterplot of Standard Deviations vs. Rates")
```

This is a scatterplot of the standard deviations of the five distributions on the x-axis versus the rates of the five distribution on the y-axis. This plot shows that as the standard deviations increase, the rates decrease. It is an exponentially distributed graph. 

```{r}
#c
plot(means, sds, main = "Scatterplot of Means vs. Standard Deviations")
```

This is a scatterplot of the means of the five distributions on the x-axis versus the standard deviations of the five distribution on the y-axis. This plot shows that as the means increase, the rates also increase. 

5. R's capacity for data and computation is large to what was available 10 years ago. a. To show this, generate 1.1 million numbers from the standard exponential distribution and store them in a vector called big.exp.draws.1. Calculate the mean and standard deviation. b. Plot a histogram of big.exp.draws.1. Does it match the function (1-e^{-x})? Should it? c. Find the mean of all of the entries in big.exp.draws.1 which are strictly greater than 1. You may need to first create a new vector to identify which elements satisfy this. d. Create a matrix, big.exp.draws.1.mat, containing the the values in big.exp.draws.1, with 1100 rows and 1000 columns. Use this matrix as the input to the hist() function and save the result to a variable of your choice. What happens to your data? e. Calculate the mean of the 371st column of big.exp.draws.1.mat. f. Now, find the means of all 1000 columns of big.exp.draws.1.mat simultaneously. Plot the histogram of column means. Explain why its shape does not match the histogram in problem 5b).

```{r}
#a
big.exp.draws.1 <- rexp(n=1100000)
meanbig <- mean(big.exp.draws.1)
sdbig <- sd(big.exp.draws.1)
```
The mean of big.exp.draws.1 is `r meanbig`.
The standard deviation of big.exp.draws.1 is `r sdbig`. 

```{r}
#b
hist(big.exp.draws.1, main = "Histogram of big.exp.draws.1")
x <- 0:10
plot(1-exp(-x), main = "Plot of function (1-e^{-x})")
plot(exp(-x), main = "Plot of function (e^{-x}")
```

No the histogram of big.exp.draws.1 does not match the function (1-e^{-x}). It should match the function e^{-x}. 

```{r}
#c
greaterthan1 <- big.exp.draws.1[which(big.exp.draws.1 > 1)] #creating a vector with values in big.exp.draws.1 greater than 1
meangreater1 <- mean(greaterthan1)
```
The mean of all of the entries in big.exp.draws.1 which are strictly greater than 1 is `r meangreater1`. 

```{r}
#d
big.exp.draws.1.mat <- matrix(big.exp.draws.1, nrow=1100, ncol=1000) #creating matrix 
histbig <- hist(big.exp.draws.1.mat, main = "Histogram of Matrix big.exp.draws.1.mat") 
```

The histogram of the matrix of big.exp.draws.1 with 1100 rows and 1000 columns looks identical to the histogram of the vector big.exp.draws.1. It is skewed right. It bins the data and because 1100 x 1000 is 1.1 million, that is why the graphs look the same.  

```{r}
#e
meancol371 <- mean(big.exp.draws.1.mat[, 371])
```
The mean of the 371st column of big.exp.draws.1.mat is `r meancol371`. 

```{r}
#f
meanallcol <- colMeans(big.exp.draws.1.mat)
hist(meanallcol, main = "Histogram of All Column Means")
```

Because we are taking the mean of the columns first, then the graph will become more normally distributed. Even if the original independent random variables are not normally distributed, as seen that the histogram of 5b is skewed right, when the mean of these values is calculated, the distribution of the average will be close to a normal distribution.  
