---
title: "pre-class-04.Rmd"
output: html_document
---

Make sure you commit this often with meaningful messages. 

### Question 1:
Using a loop, print the integers from 1 to 50. 
```{r}
for (i in 1:50)
  print(i)
```

### Question 2:
A.  Using a loop, add all the integers between 0 and 1000.
```{r}
sum = 0 #establishing local sum is equal to 0 
for (i in 0:1000){
  sum = sum + i
}
print(sum) #printing the sum of all the integers between 0 and 1000
```
B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
sum = 0 #establishing local sum is equal to 0 
for (i in seq.int(0,1000, by = 2)) {
  sum = sum + i
}
print(sum) #printing the sum of all even integers between 0 and 1000
```

C. Now, repeat A and B WITHOUT using a loop.
```{r}
sumAll <- sum(1:1000) #sum all integers between 0 and 1000
print(sumAll)
sumEven <- sum(seq.int(0,1000, by = 2)) #sum of all even integers between 0 and 1000
print(sumEven)
```

### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.
```{r}
survey.clean <- survey #setting survey.clean to survey 

for (i in 1:6) { #looping through rows (each participant) 
  for (j in 2:6 ) { #looping through columns (each question)
  if (survey[i,j] <1 | survey[i,j] >5) {
    survey.clean[i,j] = NA #setting equal to NA if integer not betwen 1 and 5 
  } else {
    survey.clean[i,j] = survey[i,j] #keeping same if integer IS between 1 and 5
}
  }
}
print(survey.clean)
```


B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.
```{r}
survey.clean$invalid.answers <- rep(0)

for (i in 1:6) {#looping through rows
  for (j in 2:6 ) {#looping through column 
    if (is.na(survey.clean[i,j]) == TRUE) { #if value is NA 
      survey.clean[i,7] = survey.clean[i,7] + 1 #add one to the new column
    }
    }
}

print(survey.clean)
```
