---
title: "pre-class-06"
output: html_document
---

Make sure you commit this often with meaningful messages.

knitr::opts_chunk$set(echo = TRUE)
1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.
```{r}
haveprefix <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
haveprefix("hello my name is amy", "hello")


deletelast <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
deletelast(c(1,0,2,3,4,9,30,2,3))

repeatlength<- function(x, y) {
  rep(y, length.out = length(x))
}

repeatlength(c(1,20,3,4), c(1,2,3,3,33))
```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?
rnorm() generates random samples from a univariate normal distribution. mvrnorm() generates random samples from a multivariate normal distribution. The argument for rnorm() includes n, mean and sd. The argument for mvrnorm() includes n, mu and sigma. They both sample from a normal distribution and they both take in n, the number of observations. To make rnorm() and mvrnorm() more consistent, I would make the names of the arguments for mvnorm() the same as the name of the arguments for rnorm().  

3. Use lapply() and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset.
```{r}
library(dplyr)
coeffVariation <- lapply(mtcars, function(i) sd(i)/mean(i))
```


4. Use vapply() to: 

a. Compute the standard deviation of every column in a numeric data frame. 
```{r}
class(df) #check the class
columnsd <- function(df) {
  vapply(df, sd, numeric(1))
}
```

b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
```{r}
class(df) #check the class 
sd_numcol <- function(df) {
  vapply(df[vapply(df, is.numeric, logical(1))], sd, numeric(1))
}
```
