---
title: "Pre-class-05"
output: html_document
---
```{r}
suppressMessages(library(foreign))
library(dplyr)
wcgs <- read.dta("https://drive.google.com/uc?export=download&id=0B8CsRLdwqzbzYWxfN3ExQllBQkU")
```

In-Class Problems
Find the files here
Save the README.md and pre-class-05.Rmd in your pre-class repository under a new folder called Week 05.
Remember to commit often. 
Standardizing a variable means subtracting the mean, and then dividing by the standard deviation. Let’s use a loop to standardize the numeric columns in the Western Collaborative Group Study. This study began in 1960 with 3154 men ages 39-59, who were employed in one of 11 California based companies. They were followed until 1969 during this time, 257 of these men developed coronary heart disease (CHD).

The data has the following variables:

WCGS has the following variables:

Name Description

id Subject identification number

age Age in years

height Height in inches

weight Weight in lbs.

sbp Systolic blood pressure in mm

dbp Diastolic blood pressure in mm Hg

chol Fasting serum cholesterol in mm

behpat Behavior

1 A1

2 A2

3 B3

4 B4

ncigs Cigarettes per day

dibpat Behavior

1 type A

2 type B

chd69 Coronary heart disease

1 Yes

0 no

typechd Type of CHD

1 myocardial infarction or death

2 silent myocardial infarction

3 angina perctoris

time169 Time of CHD event or end of follow-up

arcus Arcus senilis

0 absent

1 present

bmi Body Mass Index
Question 1: Standardize Function
A. Create a function called standardize.me() that takes a numeric vector as an argument, and returns the standardized version of the vector. 
```{r}
standardize.me <- function(x) {
  standardized <- (x - mean(x, na.rm = TRUE))/sd(x, na.rm = TRUE) #subtracting mean and dividing by sd 
  return(standardized)
}
```

B. Assign all the numeric columns of the original WCGS dataset to a new dataset called WCGS.new. 
```{r}
WCGS.new <- select_if(wcgs, is.numeric) #selects only numeric columns
```

C. Using a loop and your new function, standardize all the variables WCGS.new dataset. 
```{r}
for(i in 1:ncol(WCGS.new)) {
  WCGS.new[,i] <- standardize.me(WCGS.new[,i]) #use standardize.me function to standardize all variables of WCGS.new dataset
}
```

D. What should the mean and standard deviation of all your new standardized variables be? Test your prediction by running a loop
```{r}
#the mean should be zero and the standard deviation should be one 
for (i in 1:ncol(WCGS.new)) {
  mean1 <- round(mean(WCGS.new[,i]))
  sd1 <- round(sd(WCGS.new[,i]))
}
mean1
sd1
```


Question 2: Looping to Calculate
A. Using a loop, calculate the mean weight of the subjects separated by the type of CHD they have.
```{r}
chdtypes <- as.vector(unlist(unique(wcgs["typchd69"]))) #creates vector with each type of CHD 
means <- c(0,0,0,0) #creates vector "means" with four zeros because there are 4 different types of CHD 
df <- data.frame(chd_types, means) #creates data frame with column for type of CHD and then respective means 

for(i in 1:length(chdtypes)) {
  df[i, 2] <- mean(wcgs[wcgs$typchd69 == types[i], ]$weight) #calculates mean of each type of CHD 
}
df #output data frame/results
```


B. Now do the same thing, but now don’t use a loop
```{r}
wcgs %>%
  group_by(typchd69) %>%
  summarise(mean(weight))
```

